% =============================================================================
\section{Token Architecture}
\label{sec:architecture}
% =============================================================================

This section specifies the CORTEX token parameters, allocation structure,
vesting schedules, and smart contract architecture that form the foundation of
the PURECORTEX economy.

\subsection{Token Parameters}
\label{subsec:token-params}

Table~\ref{tab:token-params} summarizes the core token parameters.

\begin{table}[H]
\centering
\caption{CORTEX token parameters.}
\label{tab:token-params}
\begin{tabular}{@{}ll@{}}
\toprule
\textbf{Parameter} & \textbf{Value} \\
\midrule
Name & CORTEX \\
Standard & Algorand Standard Asset (ASA) \\
Total Supply & $10{,}000{,}000{,}000{,}000{,}000$ ($10^{16}$, i.e., 10 quadrillion) \\
Decimals & 6 \\
Smallest Unit & $10^{-6}$ CORTEX ($1\;\mu\text{CORTEX}$) \\
Chain & Algorand MainNet \\
Finality & Instant ($< 3.3$~s block time) \\
Transferability & Unrestricted after TGE \\
\bottomrule
\end{tabular}
\end{table}

The large total supply ($10^{16}$) with six decimal places ensures sufficient
granularity for micropayment flows in agent-to-agent transactions (see
Section~\ref{sec:composability}) while maintaining human-readable unit prices
during early-stage price discovery.

\subsection{Allocation Structure}
\label{subsec:allocation}

The total supply is partitioned into six categories as specified in
Table~\ref{tab:allocation}. The allocation reflects a community-first
philosophy: \textbf{0\% to VCs, 0\% to team/advisors} --- only a single
creator allocation of 10\%, with 90\% directed to community and ecosystem
participants.

\begin{table}[H]
\centering
\caption{CORTEX token allocation.}
\label{tab:allocation}
\begin{tabular}{@{}lrrl@{}}
\toprule
\textbf{Category} & \textbf{Percentage} & \textbf{Tokens ($\times 10^{12}$)} & \textbf{Vesting} \\
\midrule
Creator & 10\% & 1{,}000 & 10\% at TGE, 90\% daily over 180 days \\
Genesis Distribution & 31\% & 3{,}100 & Airdrop to early testnet users \& Algorand community \\
Future Emissions \& Staking & 24\% & 2{,}400 & veCORTEX rewards, 4-year halving decay \\
Liquidity & 15\% & 1{,}500 & CORTEX/ALGO DEX LP (10-year lock) + agent token LPs \\
Agent Incentives & 15\% & 1{,}500 & Graduation bonuses, composability rewards, performance \\
Assistance Fund & 5\% & 500 & Seed for buyback engine; self-sustaining from revenue \\
\midrule
\textbf{Total} & \textbf{100\%} & \textbf{10{,}000} & \\
\bottomrule
\end{tabular}
\end{table}

The Assistance Fund receives 90\% of \emph{all} protocol revenue (trading
fees, agent creation fees, x402 commissions, LP fees) and uses it to
continuously buy CORTEX from the open market and burn it. The initial 5\%
seed provides the fund with operational capital at launch; thereafter it is
entirely self-sustaining from protocol revenue. The remaining 10\% of
protocol revenue covers operations and development (see
Section~\ref{sec:dynamic-fees}).

\subsection{Vesting Formulas}
\label{subsec:vesting}

We define three distinct vesting schedules, each with a closed-form expression.

\subsubsection{Creator Vesting}

The creator allocation ($A_c = 10\%$ of total supply) vests according to:

\begin{equation}
  V_{\text{creator}}(d) = 0.10 \cdot A_c + 0.90 \cdot A_c \cdot \frac{\min(d, 180)}{180}
  \label{eq:creator-vesting}
\end{equation}

This releases 10\% of the allocation at the Token Generation Event (enabling
the creator to provide initial liquidity and demonstrate commitment) and vests
the remaining 90\% linearly over 180 days via daily unlocks. The creator's
unvested tokens are held in a smart contract escrow; premature withdrawal
attempts revert. There is no separate team or advisor allocation --- the
creator allocation is the sole insider distribution.

\subsubsection{Staking Emissions}

The staking emission pool of $R_{\text{total}} = 2.4 \times 10^{15}$ tokens
(24\% of total supply) is distributed over a four-year period with declining
annual weights:

\begin{equation}
  E(y) = R_{\text{total}} \cdot \frac{w_y}{\sum_{i=1}^{4} w_i}
  \label{eq:staking-emission}
\end{equation}

where the weight vector is $\mathbf{w} = (0.4,\; 0.3,\; 0.2,\; 0.1)$ and $y
\in \{1,2,3,4\}$ denotes the year. This yields:

\begin{align}
  E(1) &= 2.4 \times 10^{15} \times 0.4 = 9.6 \times 10^{14} \label{eq:emission-y1} \\
  E(2) &= 2.4 \times 10^{15} \times 0.3 = 7.2 \times 10^{14} \label{eq:emission-y2} \\
  E(3) &= 2.4 \times 10^{15} \times 0.2 = 4.8 \times 10^{14} \label{eq:emission-y3} \\
  E(4) &= 2.4 \times 10^{15} \times 0.1 = 2.4 \times 10^{14} \label{eq:emission-y4}
\end{align}

The declining schedule front-loads rewards to incentivize early staking while
ensuring four full years of emission-funded yield.

\subsubsection{Genesis Distribution}

The genesis distribution of $A_{\text{gen}} = 3.1 \times 10^{15}$ tokens
(31\% of total supply) is released in full at TGE via airdrop to early testnet
participants and the Algorand community. There is no vesting schedule for this
allocation --- recipients have immediate, unrestricted access:

\begin{equation}
  V_{\text{gen}}(d) = A_{\text{gen}} \quad \forall\; d \geq 0
  \label{eq:genesis-vesting}
\end{equation}

The immediate release ensures broad distribution from day one and maximizes
the decentralization of the circulating supply at launch.

\subsection{Smart Contract Architecture}
\label{subsec:contract-arch}

The PURECORTEX protocol is implemented as a system of three interacting
Algorand smart contracts, depicted schematically in
Figure~\ref{fig:contract-arch}.

\begin{figure}[H]
\centering
\begin{tikzpicture}[
  box/.style={draw, rounded corners=3pt, minimum width=3.8cm, minimum height=1.4cm,
              fill=blue!8, font=\small\sffamily, align=center},
  arrow/.style={-{Stealth[length=3mm]}, thick},
  label/.style={font=\tiny\sffamily, midway, fill=white, inner sep=1pt}
]
  % Nodes
  \node[box] (factory) at (0,0) {AgentFactory\\{\tiny ASA creation, bonding curve,}\\{\tiny graduation, vesting}};
  \node[box] (treasury) at (6,0) {Assistance Fund\\{\tiny 90\% revenue $\to$ buyback-burn,}\\{\tiny 10\% $\to$ operations}};
  \node[box] (ve) at (3,-3.5) {veCORTEX\\{\tiny Lock, decay, boost,}\\{\tiny revenue distribution}};
  \node[box, fill=green!8] (tinyman) at (9,-3.5) {Tinyman DEX\\{\tiny LP pools, locked}\\{\tiny liquidity}};
  \node[box, fill=orange!8] (governance) at (-3,-3.5) {Governance\\{\tiny Senator, proposals,}\\{\tiny constitution}};

  % Arrows
  \draw[arrow] (factory) -- node[label, above] {fees} (treasury);
  \draw[arrow] (treasury) -- node[label, right, pos=0.4] {supply reduction} (ve);
  \draw[arrow] (factory) -- node[label, left, pos=0.4] {staking emissions} (ve);
  \draw[arrow] (factory.south east) -- node[label, right, pos=0.5] {graduation LP} (tinyman);
  \draw[arrow] (treasury) -- node[label, above] {buyback} (tinyman);
  \draw[arrow] (ve) -- node[label, above] {voting weight} (governance);
  \draw[arrow] (governance) -- node[label, left, pos=0.4] {parameter updates} (factory);
\end{tikzpicture}
\caption{Smart contract architecture. Arrows indicate primary data and value
flows between contracts. Tinyman DEX and Governance are external to the core
protocol but interact through well-defined interfaces.}
\label{fig:contract-arch}
\end{figure}

\begin{description}
  \item[AgentFactory.] Handles agent creation, bonding curve pricing, token
    minting via ASA inner transactions, creator vesting escrow, and graduation
    (atomic LP formation on Tinyman). Stores per-agent state in Algorand box
    storage: current supply, base price, slope, graduation status, and creator
    vesting schedule.

  \item[SovereignTreasury (Assistance Fund).] Receives 90\% of all protocol
    revenue from trading fees, agent creation fees, x402 commissions, and LP
    fees. Executes continuous buyback-and-burn operations
    (Section~\ref{sec:buyback-burn}) by purchasing CORTEX from the open DEX
    market and sending it to the Algorand zero address. The remaining 10\% of
    protocol revenue is directed to an operations account for development and
    infrastructure costs. Seeded with 5\% of total supply at genesis;
    thereafter entirely self-sustaining from protocol revenue.

  \item[veCORTEX.] Manages token locking, computes time-decaying governance
    weight, and calculates boost multipliers for staking emissions. Lock state
    is stored in per-user box storage with fields for locked amount, lock
    duration, lock start timestamp, and accumulated emission rewards. Fee
    revenue no longer flows directly to stakers --- instead, stakers benefit
    from the continuous supply reduction executed by the Assistance Fund.
\end{description}

All three contracts interact through Algorand atomic transaction groups,
ensuring that multi-step operations (e.g., graduation: sell remaining bonding
curve supply, form LP pair, lock LP tokens, distribute graduation bonus) either
complete entirely or revert entirely. This atomicity eliminates the reentrancy
and partial-execution risks inherent in EVM-based designs.
